### 2016-04-27 manuscript/equations.Rmd

---
title: "Equations & R Code"
date: "March 22, 2016"
output:
  pdf_document:
    fig_caption: yes
---

Seasonal (within-year) dynamics are modeled as three differential equations:

\begin{align}
\frac{\text{d}N_{i}}{\text{d}\tau} &= N_{i}\epsilon_if_{i}(R)\\
\frac{\text{d}R}{\text{d}\tau} &= - \sum\limits_{i=1,2}f_{i}(R)N_{i}
\end{align}

```{r, eval=F, echo=TRUE}
## Continuous model
updateNR <- function(t, NR, parms){
  with(as.list(c(NR, parms)), {
    dN1dt = N1*eps[1]*(uptake_R(r[1], R, a[1], b[1]))
    dN2dt = N2*eps[2]*(uptake_R(r[2], R, a[2], b[2]))
    dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
    list(c(dN1dt, dN2dt, dRdt)) #output
  })
}

##  Resource uptake function (Hill function)
uptake_R <- function(r, R, a, b){
  return((r*R^a) / (b^a + R^a))
}
```

\noindent where $i$ denotes species, $D$ is the dormant (long-lived) biomass state, and $N$ is the living biomass (fast-growing, shorter-lived) state, and $\epsilon_i$ is each species' resource-to-biomass conversion efficiency. The growth rate of living biomass is a resource-dependent Hill function, $f_{i}(R) = r_{i}R^{a_{i}} / (b_{i}^{a_{i}}+R^{a_{i}})$, where *r* is a species' intrinsic growth rate and *a* and *b* define the curvature of the function. Resource depletion is equal to the sum of each species' consumption, $\sum_{i=1,2}f_{i}(R)N_{i}$.

At the beginning of each season we start with initial conditions defined as $V_{t}$, $W_{t}$, and $Z_{t}$ for the dormant state, the live state, and the resource, respectively. So for each season, eqs. 1-3 are solved given the initial conditions:
\begin{align}
  D_{i}(0) &= V_{i,t} \\
  N_{i}(0) &= W_{i,t} \\
  R(0) &= Z_{t}
\end{align}

\noindent The consumers transition between $N$ and $D$ instantaneously between years. So, at the yearly transition:

\begin{align}
  V_{i,t+1} &= \alpha_i N_{i}(T^-) + D_{i}(T^-)(1-\gamma_{i,t})(1-\eta_i) \\
  W_{i,t+1} &= \beta_i(1-\alpha_i)N_{i}(T^-) + \gamma_{i,t}[D_{i}(T^-) + \alpha_i N_{i}(T^-)] (1-\eta_i) \\
  Z_{t+1} &= \theta_i (1-\alpha_i)N_i(T^-) + \nu R(T^-) + R(T^+)
\end{align}

```{r, eval=F, echo=T}
## Discrete model
update_DNR <- function(t,DNR,gammas,alpha1,alpha2,eta1,eta2,beta1,beta2,theta1,theta2,nu){
  with (as.list(DNR),{
    g1 <- gammas[1]
    g2 <- gammas[2]
    D1 <- alpha1*N1 + D1*(1-g1)*(1-eta1)
    D2 <- alpha2*N2 + D2*(1-g2)*(1-eta2)
    N1 <- beta1*(1-alpha1)*N1 + g1*(D1+(alpha1*N1))*(1-eta1)
    N2 <- beta2*(1-alpha2)*N2 + g2*(D2+(alpha2*N2))*(1-eta2)
    R <- theta1*(1-alpha1)*N1 + theta2*(1-alpha2)*N2 + nu*R + Rvector[t]
    return(c(D1, D2, N1, N2, R))
  })
}
```

\noindent where $D(T^-)$, $N(T^-)$, and $R(T^-)$ are the densities of each state at the end of the year and $g$ is a time-fluctuating activation rate that regulates how much dormant biomass is converted to growing-season live biomass each year. $R(T^+)$ is a randomly generated resource pulse from a log-normal distribution with mean $R_\mu$ and variance $R_{\sigma^2}$. Parameter notations and definitions are in table 1.

\newpage{}

\begin{center}
\begin{table}
\caption{Definition of model states and parameters. `Constant' parameters do not fluctuate during a simulation, while `variable' parameters do fluctuate through time in a given simulation.}
\begin{tabular}{l l}
\hline
Parameter & Definition \\
\hline
$N$ & live biomass \\
$D$ & dormant biomass \\
$R$ & resource supply \\
$\alpha$ & allocation fraction of live biomass to seed bank (constant) \\
$\beta$ & live state survivorship (constant) \\
$\eta$ & dormant state mortality rate (constant) \\
$\epsilon$ & resource-to-biomass efficiency (constant) \\
$\gamma$ & dormant-to-live biomass transition fraction (variable) \\ 
$\nu$ & resource carry-over fraction (constant) \\
$\theta$ & resource recycling fraction (constant) \\
$a$ & Hill function power parameter (constant) \\
$b$ & Hill function parameter (constant) \\
$\rho$ & correlation of species' responses transition cue (constant) \\
$\sigma^2_{E}$ & variance of transition cue (constant) \\
$R_{\mu}$ & mean resource pulse, on log scale (constant) \\
$R_{\sigma^2}$ & variance of annual resource supply, on log scale (variable) \\

\hline
\end{tabular}
\end{table}
\end{center}

## Full R Code for Simulations
```{r eval=T, echo=T, fig.cap="Abundance of both species over 800 seasons. Species have identical resource uptake rates, but differ in their conversion efficiency. This 2 species community represents and annual plant community (no biomass carry-over) with no resource recycling."}
####
#### Initial Conditions, Global Variables, and Parameters ----------------------
####
seasons <- 1000                  # number of seasons to simulate
seasons_to_exclude <- 200        # initial seasons to exclude from plots
days_to_track <- 20              # number of days to simulate in odSolve
DNR <- c(D=c(1,1),N=c(1,1),R=10) # initial conditions
Rmu <- 3                         # mean resource pulse (on log scale)
Rsd_annual <- 0                  # std dev of resource pulses (on log scale)
sigE <- 2                        # environmental cue variance
rho <- 0                         # environmental cue correlation between species

# Within-season parameters
parms <- list(
  r = c(5,5),                    # max growth rate for each species
  a = c(5,5),                    # rate parameter for Hill function 
  b = c(20,20),                  # shape parameter for Hill function
  eps = c(0.2,0.2)               # resource-to-biomass efficiency
)

# End-of-season transition parameters
alpha1 <- 0.50                   # live-to-dormant biomass fraction; spp1
alpha2 <- 0.49                   # live-to-dormant biomass fraction; spp2
beta1 <- 0                       # adult survivorship; spp1 (0 if annual, >0 if perennial)
beta2 <- 0                       # adult survivorship; spp2 (0 if annual, >0 if perennial)
eta1 <- 0.1                      # dormant mortality; spp1
eta2 <- 0.1                      # dormant mortality; spp2
theta1 <- 0                      # resource recycling fraction; spp1
theta2 <- 0                      # resource recycling fraction; spp2
nu <- 0                          # resource carry-over fraction

####
#### Load relevant libraries ----------------------------------------
####
library('deSolve')
library('mvtnorm')

####
#### Model functions -------------------------------------------------
####
## Continuous model
updateNR <- function(t, NR, parms){
  with(as.list(c(NR, parms)), {
    dN1dt = N1*eps[1]*(uptake_R(r[1], R, a[1], b[1]))
    dN2dt = N2*eps[2]*(uptake_R(r[2], R, a[2], b[2]))
    dRdt = -1 * (dN1dt/eps[1] + dN2dt/eps[2])
    list(c(dN1dt, dN2dt, dRdt)) #output
  })
}

##  Resource uptake function (Hill function)
uptake_R <- function(r, R, a, b){
  return((r*R^a) / (b^a + R^a))
}

## Discrete model
update_DNR <- function(t,DNR,gammas,alpha1,alpha2,eta1,eta2,beta1,beta2,theta1,theta2,nu){
  with (as.list(DNR),{
    g1 <- gammas[1]
    g2 <- gammas[2]
    D1 <- alpha1*N1 + D1*(1-g1)*(1-eta1)
    D2 <- alpha2*N2 + D2*(1-g2)*(1-eta2)
    N1 <- beta1*(1-alpha1)*N1 + g1*(D1+(alpha1*N1))*(1-eta1)
    N2 <- beta2*(1-alpha2)*N2 + g2*(D2+(alpha2*N2))*(1-eta2)
    R <- theta1*(1-alpha1)*N1 + theta2*(1-alpha2)*N2 + nu*R + Rvector[t]
    return(c(D1, D2, N1, N2, R))
  })
}

####
#### Simulate model -----------------------------------------------------
####
days <- c(1:days_to_track)

# Get "germination" fractions for each year
getG <- function(sigE, rho, nTime){
  varcov <- matrix(c(sigE, rho*sigE, rho*sigE, sigE), 2, 2)
  e <- rmvnorm(n = nTime, mean = c(0,0), sigma = varcov)
  g <- exp(e) / (1+exp(e))
  return(g)
}

# Loop over seasons
nmsDNR <- names(DNR)
NR <- DNR[c("N1", "N2", "R")] 
nmsNR <- names(NR)
gVec <- getG(sigE = sigE, rho = rho, nTime = seasons)
Rvector <- rlnorm(seasons, Rmu, Rsd_annual)
saved_outs <- matrix(ncol=5, nrow=seasons)
for(season_now in 1:seasons){
  output <- ode(y = NR, times=days,
                func = updateNR, parms = parms)
  NR <- output[nrow(output),nmsNR]
  DNR <- c(DNR[c("D1","D2")], NR)
  saved_outs[season_now,] <- DNR
  names(DNR) <- nmsDNR
  DNR <- update_DNR(season_now, DNR, gVec[season_now,],
                    alpha1=alpha1,alpha2=alpha2,
                    eta1=eta1,eta2=eta2,
                    beta1=beta1,beta2=beta2,
                    theta1=theta1,theta2=theta2,nu=nu)
  names(DNR) <- nmsDNR
  NR <- DNR[c("N1", "N2", "R")] 
  names(NR) <- nmsNR
}
matplot(saved_outs[101:seasons,c(3:4)], type="l", 
        xlab="Season", ylab="Abundance",
        main=expression(paste(alpha[1],"= 0.5; ", alpha[2],"=0.49; ", sigma[E],"=0.2")))
```