### 2016-11-02 manuscript/coexistence-stability_SI.tex
adds SI mateerial; cleans up figures

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
\newcommand{\NormalTok}[1]{{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{todonotes}
\onehalfspacing
\usepackage{rotating}
\usepackage{color, soul}
\usepackage[font={small},labelfont={sf,bf},labelsep=space]{caption}
\usepackage{tikz}

\begin{document}

\maketitle

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{S1-\arabic{figure}}  
\renewcommand\thetable{S1-\arabic{table}}  
\renewcommand\thesection{Section S1.\arabic{section}}

\begin{center}
\textbf{\Large{Supporting Information Appendix S1}} \\
A.T. Tredennick, P.B. Adler, \& F.R. Adler, ``How species coexistence affects...'' \\
\emph{Journal Title}
\end{center}

\section{R Code for Consumer-Resource Model}

Below is the R code for our model function, which is represented
mathematically in the main text in Equations 1-4. The same code, along
with all the code to reproduce our results, has been archived on
Figshare (link) and is available on GitHub
(\url{http://github.com/atredennick/Coexistence-Stability/releases}).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{simulate_model <-}\StringTok{ }\NormalTok{function(seasons, days_to_track, Rmu, }
                           \NormalTok{Rsd_annual, sigE, rho, }
                           \NormalTok{alpha1, alpha2, alpha3, alpha4,}
                           \NormalTok{beta1, beta2, beta3, beta4,}
                           \NormalTok{eta1, eta2, eta3, eta4,}
                           \NormalTok{theta1, theta2, theta3, theta4,}
                           \NormalTok{nu, r1, r2, r3, r4,}
                           \NormalTok{a1, a2, a3, a4,}
                           \NormalTok{b1, b2, b3, b4,}
                           \NormalTok{eps1, eps2, eps3, eps4,}
                           \NormalTok{D1, D2, D3, D4,}
                           \NormalTok{N1, N2, N3, N4, R) \{}

  \KeywordTok{require}\NormalTok{(}\StringTok{'deSolve'}\NormalTok{) }\CommentTok{# for solving continuous differential equations}
  \KeywordTok{require}\NormalTok{(}\StringTok{'mvtnorm'}\NormalTok{) }\CommentTok{# for multivariate normal distribution functions}

  \NormalTok{##  Assign parameter values to appropriate lists}
  \NormalTok{DNR <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DataTypeTok{D=}\KeywordTok{c}\NormalTok{(D1,D2,D3,D4),   }\CommentTok{# initial dormant state abundance}
           \DataTypeTok{N=}\KeywordTok{c}\NormalTok{(N1,N2,N3,N4),   }\CommentTok{# initial live state abundance}
           \DataTypeTok{R=}\NormalTok{R)                }\CommentTok{# initial resource level}

  \NormalTok{parms <-}\StringTok{ }\KeywordTok{list} \NormalTok{(}
    \DataTypeTok{r   =} \KeywordTok{c}\NormalTok{(r1,r2,r3,r4),          }\CommentTok{# max growth rate for each species}
    \DataTypeTok{a   =} \KeywordTok{c}\NormalTok{(a1,a2,a3,a4),          }\CommentTok{# rate parameter for Hill function }
    \DataTypeTok{b   =} \KeywordTok{c}\NormalTok{(b1,b2,b3,b4),          }\CommentTok{# shape parameter for Hill function}
    \DataTypeTok{eps =} \KeywordTok{c}\NormalTok{(eps1,eps2,eps3,eps4)   }\CommentTok{# resource-to-biomass efficiency}
  \NormalTok{)}

  \NormalTok{####}
  \NormalTok{####  Sub-Model functions ----------------------------------------------------}
  \NormalTok{####}
  \NormalTok{## Continuous model (Equations 1-2)}
  \NormalTok{updateNR <-}\StringTok{ }\NormalTok{function(t, NR, parms)\{}
    \KeywordTok{with}\NormalTok{(}\KeywordTok{as.list}\NormalTok{(}\KeywordTok{c}\NormalTok{(NR, parms)), \{}
      \NormalTok{dN1dt =}\StringTok{ }\NormalTok{N1*eps[}\DecValTok{1}\NormalTok{]*(}\KeywordTok{uptake_R}\NormalTok{(r[}\DecValTok{1}\NormalTok{], R, a[}\DecValTok{1}\NormalTok{], b[}\DecValTok{1}\NormalTok{]))}
      \NormalTok{dN2dt =}\StringTok{ }\NormalTok{N2*eps[}\DecValTok{2}\NormalTok{]*(}\KeywordTok{uptake_R}\NormalTok{(r[}\DecValTok{2}\NormalTok{], R, a[}\DecValTok{2}\NormalTok{], b[}\DecValTok{2}\NormalTok{]))}
      \NormalTok{dN3dt =}\StringTok{ }\NormalTok{N3*eps[}\DecValTok{3}\NormalTok{]*(}\KeywordTok{uptake_R}\NormalTok{(r[}\DecValTok{3}\NormalTok{], R, a[}\DecValTok{3}\NormalTok{], b[}\DecValTok{3}\NormalTok{]))}
      \NormalTok{dN4dt =}\StringTok{ }\NormalTok{N4*eps[}\DecValTok{4}\NormalTok{]*(}\KeywordTok{uptake_R}\NormalTok{(r[}\DecValTok{4}\NormalTok{], R, a[}\DecValTok{4}\NormalTok{], b[}\DecValTok{4}\NormalTok{]))}
      \NormalTok{dRdt  =}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{(dN1dt/eps[}\DecValTok{1}\NormalTok{] +}\StringTok{ }\NormalTok{dN2dt/eps[}\DecValTok{2}\NormalTok{] +}\StringTok{ }\NormalTok{dN3dt/eps[}\DecValTok{3}\NormalTok{] +}\StringTok{ }\NormalTok{dN4dt/eps[}\DecValTok{4}\NormalTok{])}
      \KeywordTok{list}\NormalTok{(}\KeywordTok{c}\NormalTok{(dN1dt, dN2dt, dN3dt, dN4dt, dRdt)) }\CommentTok{# output as list}
    \NormalTok{\})}
  \NormalTok{\} }\CommentTok{# end continuous function}

  \NormalTok{## Discrete model (Equations 3-4)}
  \NormalTok{update_DNR <-}\StringTok{ }\NormalTok{function(t, DNR, gammas,}
                         \NormalTok{alpha1, alpha2, alpha3, alpha4,}
                         \NormalTok{eta1, eta2, eta3, eta4,}
                         \NormalTok{beta1, beta2, beta3, beta4,}
                         \NormalTok{theta1, theta2, theta3, theta4, nu) \{}
    \KeywordTok{with} \NormalTok{(}\KeywordTok{as.list}\NormalTok{(DNR),\{}
      \NormalTok{g1    <-}\StringTok{ }\NormalTok{gammas[}\DecValTok{1}\NormalTok{]}
      \NormalTok{g2    <-}\StringTok{ }\NormalTok{gammas[}\DecValTok{2}\NormalTok{]}
      \NormalTok{g3    <-}\StringTok{ }\NormalTok{gammas[}\DecValTok{3}\NormalTok{]}
      \NormalTok{g4    <-}\StringTok{ }\NormalTok{gammas[}\DecValTok{4}\NormalTok{]}
      \NormalTok{D1new <-}\StringTok{ }\NormalTok{alpha1*N1 +}\StringTok{ }\NormalTok{D1*(}\DecValTok{1}\NormalTok{-g1)*(}\DecValTok{1}\NormalTok{-eta1)}
      \NormalTok{D2new <-}\StringTok{ }\NormalTok{alpha2*N2 +}\StringTok{ }\NormalTok{D2*(}\DecValTok{1}\NormalTok{-g2)*(}\DecValTok{1}\NormalTok{-eta2)}
      \NormalTok{D3new <-}\StringTok{ }\NormalTok{alpha3*N3 +}\StringTok{ }\NormalTok{D3*(}\DecValTok{1}\NormalTok{-g3)*(}\DecValTok{1}\NormalTok{-eta3)}
      \NormalTok{D4new <-}\StringTok{ }\NormalTok{alpha4*N4 +}\StringTok{ }\NormalTok{D4*(}\DecValTok{1}\NormalTok{-g4)*(}\DecValTok{1}\NormalTok{-eta4)}
      \NormalTok{N1new <-}\StringTok{ }\NormalTok{beta1*(}\DecValTok{1}\NormalTok{-alpha1)*N1 +}\StringTok{ }\NormalTok{g1*(D1+(alpha1*N1))*(}\DecValTok{1}\NormalTok{-eta1)}
      \NormalTok{N2new <-}\StringTok{ }\NormalTok{beta2*(}\DecValTok{1}\NormalTok{-alpha2)*N2 +}\StringTok{ }\NormalTok{g2*(D2+(alpha2*N2))*(}\DecValTok{1}\NormalTok{-eta2)}
      \NormalTok{N3new <-}\StringTok{ }\NormalTok{beta3*(}\DecValTok{1}\NormalTok{-alpha3)*N3 +}\StringTok{ }\NormalTok{g3*(D3+(alpha3*N3))*(}\DecValTok{1}\NormalTok{-eta3)}
      \NormalTok{N4new <-}\StringTok{ }\NormalTok{beta4*(}\DecValTok{1}\NormalTok{-alpha4)*N4 +}\StringTok{ }\NormalTok{g4*(D4+(alpha4*N4))*(}\DecValTok{1}\NormalTok{-eta4)}
      \NormalTok{Rnew  <-}\StringTok{ }\NormalTok{theta1*(}\DecValTok{1}\NormalTok{-alpha1)*N1 +}\StringTok{ }\NormalTok{theta2*(}\DecValTok{1}\NormalTok{-alpha2)*N2 +}\StringTok{ }
\StringTok{               }\NormalTok{theta3*(}\DecValTok{1}\NormalTok{-alpha3)*N3 +}\StringTok{ }\NormalTok{theta4*(}\DecValTok{1}\NormalTok{-alpha4)*N4 +}\StringTok{ }
\StringTok{               }\NormalTok{nu*R +}\StringTok{ }\NormalTok{Rvector[t]}
      \KeywordTok{return}\NormalTok{(}\KeywordTok{c}\NormalTok{(D1new, D2new, D3new, D4new, N1new, N2new, N3new, N4new, Rnew))}
    \NormalTok{\})}
  \NormalTok{\}}

  \NormalTok{##  Resource uptake function (Hill function)}
  \NormalTok{uptake_R <-}\StringTok{ }\NormalTok{function(r, R, a, b) \{}
    \KeywordTok{return}\NormalTok{((r*R^a) /}\StringTok{ }\NormalTok{(b^a +}\StringTok{ }\NormalTok{R^a))}
  \NormalTok{\}}

  \NormalTok{##  Generate germination fractions}
  \NormalTok{getG <-}\StringTok{ }\NormalTok{function(sigE, rho, nTime, num_spp) \{}
    \NormalTok{varcov       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{rep}\NormalTok{(rho*sigE,num_spp*}\DecValTok{2}\NormalTok{), num_spp, num_spp)}
    \KeywordTok{diag}\NormalTok{(varcov) <-}\StringTok{ }\NormalTok{sigE}
    \NormalTok{if(sigE >}\StringTok{ }\DecValTok{0}\NormalTok{) \{ varcov <-}\StringTok{ }\NormalTok{Matrix::}\KeywordTok{nearPD}\NormalTok{(varcov)$mat \} }\CommentTok{# coerce matrix to PD }
    \NormalTok{varcov <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(varcov)}
    \NormalTok{e      <-}\StringTok{ }\KeywordTok{rmvnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nTime, }\DataTypeTok{mean =} \KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,num_spp), }\DataTypeTok{sigma =} \NormalTok{varcov)}
    \NormalTok{g      <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(e) /}\StringTok{ }\NormalTok{(}\DecValTok{1}\NormalTok{+}\KeywordTok{exp}\NormalTok{(e))}
    \KeywordTok{return}\NormalTok{(g)}
  \NormalTok{\}}

  \NormalTok{####}
  \NormalTok{#### Simulate model -----------------------------------------------------}
  \NormalTok{####}
  \NormalTok{days           <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{:days_to_track)}
  \NormalTok{num_spp        <-}\StringTok{ }\KeywordTok{length}\NormalTok{(parms$r)}
  \NormalTok{nmsDNR         <-}\StringTok{ }\KeywordTok{names}\NormalTok{(DNR)}
  \NormalTok{dormants       <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\KeywordTok{names}\NormalTok{(DNR))}
  \NormalTok{NR             <-}\StringTok{ }\NormalTok{DNR[-dormants] }
  \NormalTok{nmsNR          <-}\StringTok{ }\KeywordTok{names}\NormalTok{(NR)}
  \NormalTok{gVec           <-}\StringTok{ }\KeywordTok{getG}\NormalTok{(}\DataTypeTok{sigE =} \NormalTok{sigE, }\DataTypeTok{rho =} \NormalTok{rho, }\DataTypeTok{nTime =} \NormalTok{seasons, }\DataTypeTok{num_spp =} \NormalTok{num_spp)}
  \NormalTok{Rvector        <-}\StringTok{ }\KeywordTok{rlnorm}\NormalTok{(seasons, Rmu, Rsd_annual)}
  \NormalTok{saved_outs     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{ncol=}\KeywordTok{length}\NormalTok{(DNR), }\DataTypeTok{nrow=}\NormalTok{seasons}\DecValTok{+1}\NormalTok{)}
  \NormalTok{saved_outs[}\DecValTok{1}\NormalTok{,] <-}\StringTok{ }\NormalTok{DNR }

  \NormalTok{##  Loop over seasons}
  \NormalTok{for(season_now in }\DecValTok{1}\NormalTok{:seasons) \{}
    \CommentTok{# Simulate continuous growing  season}
    \NormalTok{output   <-}\StringTok{ }\KeywordTok{ode}\NormalTok{(}\DataTypeTok{y =} \NormalTok{NR, }\DataTypeTok{times=}\NormalTok{days, }\DataTypeTok{func =} \NormalTok{updateNR, }\DataTypeTok{parms =} \NormalTok{parms)}
    \NormalTok{NR       <-}\StringTok{ }\NormalTok{output[}\KeywordTok{nrow}\NormalTok{(output),nmsNR]}
    \NormalTok{dormants <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"D"}\NormalTok{, }\KeywordTok{names}\NormalTok{(DNR)) }
    \NormalTok{DNR      <-}\StringTok{ }\KeywordTok{c}\NormalTok{(DNR[dormants], NR)}

    \CommentTok{# Save end of season biomasses, before discrete transitions}
    \NormalTok{saved_outs[season_now}\DecValTok{+1}\NormalTok{,] <-}\StringTok{ }\NormalTok{DNR}

    \KeywordTok{names}\NormalTok{(DNR) <-}\StringTok{ }\NormalTok{nmsDNR}
    \NormalTok{DNR <-}\StringTok{ }\KeywordTok{update_DNR}\NormalTok{(season_now, DNR, gVec[season_now,],}
                      \DataTypeTok{alpha1 =} \NormalTok{alpha1, }\DataTypeTok{alpha2 =} \NormalTok{alpha2, }
                      \DataTypeTok{alpha3 =} \NormalTok{alpha3, }\DataTypeTok{alpha4 =} \NormalTok{alpha4,}
                      \DataTypeTok{eta1 =} \NormalTok{eta1, }\DataTypeTok{eta2 =} \NormalTok{eta2, }\DataTypeTok{eta3 =} \NormalTok{eta3, }\DataTypeTok{eta4 =} \NormalTok{eta4,}
                      \DataTypeTok{beta1 =} \NormalTok{beta1, }\DataTypeTok{beta2 =} \NormalTok{beta2, }
                      \DataTypeTok{beta3 =} \NormalTok{beta3, }\DataTypeTok{beta4 =} \NormalTok{beta4,}
                      \DataTypeTok{theta1 =} \NormalTok{theta1, }\DataTypeTok{theta2 =} \NormalTok{theta2, }
                      \DataTypeTok{theta3 =} \NormalTok{theta3, }\DataTypeTok{theta4 =} \NormalTok{theta4,}
                      \DataTypeTok{nu=}\NormalTok{nu)}

    \KeywordTok{names}\NormalTok{(DNR) <-}\StringTok{ }\NormalTok{nmsDNR}
    \NormalTok{NR         <-}\StringTok{ }\NormalTok{DNR[-dormants]  }
    \KeywordTok{names}\NormalTok{(NR)  <-}\StringTok{ }\NormalTok{nmsNR}
  \NormalTok{\} }\CommentTok{# next season}

  \KeywordTok{return}\NormalTok{(saved_outs)}

\NormalTok{\} ##  End simulation function}
\end{Highlighting}
\end{Shaded}

\newpage{}

\section{Additional Figures}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/fourspp_Ruptake_relnonlin.png}
  \caption{Resource uptake curves for each species (represented by different colors) as used in relative nonlinearity simulations.}
\end{figure}

\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6.5in]{./components/SI_invasion_factorials.png}
  \caption{Variability of community biomass and invasion growth rates of the inferior competitor under different parameter combinations. Points are mean values from 10,000 growing seasons and lines are linear fits to show trends. In \textbf{Storage Effect} plots, resource supply is held constant between growing seasons, whereas resource supply varies each year in \textbf{Relative Nonlinearity} simulations.}
\end{figure}

\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{./components/storage_effect_div+envar_varycomp_loglog_slopes.png}
  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_E$) at different levels of realized species richness from storage effect simulations. The slopes come from linear models fit to log-transformed versions of Figure 3 in the main text. For these simulations, ``symmetric competion'' (\tikzcircle{1.5pt}) refers to similar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.495, 0.49, 0.485]$ for the four species), and ``asymmetric competition'' (\tikzcircle[fill=blue]{1.5pt}) refers to more dissimilar live-to-dormant biomass allocation fractions ($\boldsymbol{\alpha} = [0.5, 0.49, 0.48, 0.47]$ for the four species).}
\end{figure}

\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{./components/relative_nonlinearity_div+envar_loglog_slopes.png}
  \caption{Slopes of linear fits for the relationship between log(\emph{CV}) and log($\sigma_R$) at different levels of realized species richness from relative nonlinearity simulations. The slopes come from linear models fit to log-transformed versions of Figure 4 in the main text.}
\end{figure}

\end{document}